\begin{abstract}
A mask can achieve high Dice overlap yet contain a small component dangerously
far from the anatomy. It may therefore be safe under a regional criterion and
unsafe under a worst-boundary criterion: \textbf{selective segmentation cannot rely on a
metric-agnostic confidence score}. A single probability map also supplies only
pixel marginals, not the joint mask law required by nonlinear deployment costs.
We address both gaps by fixing the deployed mask and metric-defined cost, then declaring a
shared-threshold level-set law as a \emph{working posterior}. Its nested
candidates provide a tractable boundary-sensitivity probe, and their cost against the fixed
action yields a one-dimensional risk integral. Our theory separates numerical
quadrature from posterior discrepancy and propagates both to area under the
risk--coverage curve (AURC) regret. Full Hausdorff distance (HD), evaluated in
image-diagonal-normalized coordinates with a fixed one-sided-empty penalty, is
the flagship criterion: its surface geometry gives a constant-one Wasserstein
posterior-discrepancy bound, conditional on closeness in that geometry. HD95 replaces the maximum by the 95th percentile of pooled bidirectional surface distances and
is a robust nonmetric extension. Dice provides a regional contrast whose exact
algorithm and count-dispersion bound explain why level-set Dice and Soft Dice
Confidence (SDC) can behave similarly. Across five binary datasets and 16 frozen conditions, the
32-node HD score lowers HD-risk AURC relative to Dice-M32 in all ten target
conditions, with every pointwise interval excluding zero; a boundary-indexed
score is best among the displayed methods in nine. A separately locked
seven-condition SegFormer-B2/Diverse Unsupervised Saliency (DUTS) extension
reproduces the HD-over-Dice result in all seven conditions, again with every
interval excluding zero. The HD--HD95 ordering is
smaller and sensitive to checkpoint and deployment threshold. A prespecified
marginal-preserving spatial copula produces large Dice gains in some datasets
but wins only 5/10 conditions, failing its directional gate. \textbf{Reliable
selection requires a metric-specific risk target and an explicit, auditable
working-posterior geometry.}
\end{abstract}
