\section{Introduction}
\label{sec:introduction}

A segmentation can overlap almost perfectly with the annotation yet contain a
small component far from the anatomy. Dice may barely change, full Hausdorff
distance (HD) calls the error catastrophic, and its 95th-percentile variant
(HD95) may suppress it when fewer than roughly five percent of pooled surface
distances are extreme. \textbf{The same deployed mask can therefore be safe by
one evaluation metric and dangerous by another.}

This exposes a risk-alignment gap. Most image-level confidence scores compress
one probability map by a mean, entropy, patch, or overlap statistic, then use
the resulting scalar for every failure definition. Yet selective deployment
needs a sharper target: \textbf{the conditional risk of the fixed action under
the metric that actually defines harm}. Pixel marginals alone do not identify
that risk for nonlinear region or boundary criteria because they do not specify
how mask pixels vary jointly.

We separate the objects that are often conflated. An evaluation metric induces
a per-case deployment cost \(L\); its conditional expectation
\(\rho_L^\star(x)\) is deployment risk; and the ideal confidence is
\(C_L^\star(x)=-\rho_L^\star(x)\). We call a score \textbf{risk-aligned} when it
targets this quantity rather than generic uncertainty or posterior spread.

The risk remains unavailable from one probability map, so we make a second
choice explicit. We freeze the deployed hard mask, draw a shared threshold
through its probability map to generate coherent nested candidates, and compare
every candidate with that same action under \(L\). Averaging the costs estimates
working-posterior risk. The shared-threshold law preserves the probability map
as a coverage function \citep{goodman,molchanov}; it is a transparent
\emph{working posterior}, not a claim that the unknown annotation posterior is
nested.

Throughout, HD and HD95 are computed after scaling each image coordinate system
to unit diagonal; this avoids confounding boundary risk with the widely varying
native image sizes. Full HD is the flagship criterion. It records the worst
relative surface displacement and penalizes a one-sided empty mask.
Here coherent level-set motion and the deployment cost share surface geometry,
yielding a constant-one Wasserstein posterior bound. We then propagate score
error to area under the risk--coverage curve (AURC) regret. HD95 replaces the maximum with the 95th percentile of pooled
bidirectional surface distances; it is a robust practical endpoint, but its
percentile operation is nonmetric. Dice supplies the regional contrast, while
Soft Dice Confidence (SDC) reveals why two differently derived count-oriented
one-map scores can remain close.

We \textbf{freeze every probability map and deployed action}, then change only
the confidence score and evaluation risk. The primary question is whether
boundary-aligned confidence improves selection under worst-boundary risk. We
next compare full and robust boundary criteria, then use Dice to expose where
the shared-threshold coupling is less informative. The final absolute-result
display spans six datasets and three target-adapted model families with a
complete \(3\times3\) score--risk design; the original five-dataset and
seven-condition confirmatory families remain separately locked.

\subsection{Related Work}
\label{sec:related}

\noindent\textbf{Selective prediction and failure detection.}
Selective classification formalizes abstention through the risk--coverage
tradeoff: a system accepts high-confidence inputs and pays task loss only on the
accepted set \citep{geifman2017selective,selectivenet,feng}. Subsequent work
emphasizes that apparent progress in failure detection can depend on the score,
error definition, and summary measure used for evaluation \citep{jaeger}. This
issue is sharper in segmentation. Classification normally fixes one zero--one
error, whereas a mask can be judged by overlap, worst boundary displacement, or
a robust surface percentile. We retain the risk--coverage and AURC formalism,
but align both quantities with the deployed segmentation risk. The resulting
question is \textbf{not merely whether a prediction is uncertain}; it is whether the
fixed deployed mask is risky under a specified notion of geometric harm.

\noindent\textbf{Segmentation quality control.}
Existing quality-control methods span several information budgets. Reverse
Classification Accuracy (RCA) fits a reverse model for each test prediction and
evaluates it against an annotated reference set \citep{valindria2017rca}.
Learned quality estimators may regress observed mask quality from validation
labels or intermediate features. Validation of Uncertainty Estimation for
Segmentation (ValUES) systematizes uncertainty evaluation in
semantic segmentation, while medical-segmentation benchmarking shows that the
conversion from pixel uncertainty to an image-level failure score can matter as
much as the underlying uncertainty estimator \citep{values,zenk}. These are
important deployment strategies, but an atlas, labelled calibration set, learned
quality head, or repeated forward passes changes what information is available.
Our primary study deliberately fixes a narrower budget: \textbf{one probability map,
one hard action, and no fitted quality predictor}. RCA and learned regressors are
therefore related alternatives rather than matched baselines.

\noindent\textbf{Aggregating one probability map.}
Even at the single-map budget, spatial aggregation is not unique. Global mean
entropy and mean maximum probability discard where uncertainty occurs. Recent
spatially aware schemes instead select a foreground-sized high-uncertainty
fraction or the most uncertain local patch \citep{guarino2026aggregation}; the
median--minimum--maximum confidence (MMMC) statistic summarizes extrema and the
median of pixel uncertainty \citep{holder2021uncertainty}. We include the
quantile foreground-ratio score (QFR-Entropy), the rank-equivalent patch-level
mean/aggregation score (PLM/PLA-10-Entropy), MMMC-Entropy, foreground entropy,
and the global summaries
as fixed references. Their common object is a \textbf{scalar aggregation of pixelwise
uncertainty}. Our score instead generates candidate masks, compares each candidate
with the same deployed action under \(L\), and aggregates the resulting
structured costs. This difference lets a boundary score retain the location and
extent of an uncertain surface excursion rather than only its entropy mass.

\noindent\textbf{Sampling agreement and threshold stability.}
Monte Carlo dropout and deep ensembles obtain multiple predictive maps through
additional stochastic or independently trained forward passes
\citep{gal2016dropout,lakshminarayanan2017ensembles}. Bayesian QuickNAT uses
agreement across dropout segmentations, including overlap-based summaries, for
structure-wise quality control \citep{roy2019bayesianquicknat}. A different
single-map stability idea appears in the Segment Anything Model (SAM), which
compares masks obtained after a
small shift of its binarization cutoff \citep{kirillov2023segmentanything}.
Our level-set expectation is related to both ideas but is \textbf{identical to neither}:
it integrates an explicitly declared continuum law over thresholds rather than
using two nearby masks or treating model members as annotation samples. We
include threshold intersection-over-union (threshold-IoU) at the single-map
budget and separately label the
three-checkpoint empirical posterior at its higher inference budget. Model
disagreement can test support and ranking mechanisms, but it does not by itself
identify human annotation variability.

\noindent\textbf{Random sets and boundary geometry.}
Coverage functions and random level sets have a long history in fuzzy-set and
random-set theory \citep{goodman,molchanov}. Shared-level constructions also
appear in probabilistic surface extraction and excursion-set analysis
\citep{poethkow,bolin}. We use the shared threshold as a \textbf{transparent comonotone
coupling} of pixel marginals: it produces coherent nested masks and makes the
otherwise implicit joint assumption auditable. Boundary-aware objectives and
metrics were developed precisely because overlap can hide spatially remote
errors \citep{karimi2020}. Our HD instantiation connects these two strands:
level sets supply a simple boundary-motion posterior, while full HD in
unit-diagonal coordinates supplies the matching mask geometry. Pooled HD95 remains a
practical robust endpoint, but its percentile operation prevents the same metric
argument.

\noindent\textbf{The closest metric-aligned predecessor.}
Soft Dice Confidence (SDC) directly aligns a one-map score with Dice by computing
soft Dice between \(p\) and \(\Yhat\), together with a Dice-specific
approximation analysis \citep{sdc}. It is our closest predecessor and a strong
baseline. \textbf{The distinction is probabilistic rather than cosmetic}: SDC is a ratio
of expected overlap and size counts, whereas level-set Dice is the expectation
of their random ratio under a declared coupling. Our count-dispersion theorem
explains when those quantities are close, and the negative coupling ablations
show why neither score consistently dominates in practice. The broader framework is not a
claim that level sets dominate SDC for regional risk; it uses the same
risk-alignment principle beyond Dice, with its clearest advantage under Hausdorff
boundary risk.

\subsection{Our Contributions}

We contribute: (i) risk-aligned confidence and a shared-threshold estimator for
any bounded metric-induced deployment cost; (ii) an HD flagship specialization linking
coherent boundary motion to a constant-one surface-Wasserstein bound; (iii) a
complete error chain from quadrature and posterior discrepancy to AURC regret,
with HD95 and Dice specializations, Exact Dice, and an SDC count-dispersion
bound; and (iv) a locked six-dataset evaluation with full cross-risk evaluation,
single-map baselines, paired uncertainty, action and checkpoint stress tests,
separately labelled higher-budget posterior diagnostics, and a predeclared
SegFormer/DUTS architecture--domain extension. The original five-dataset
confirmatory design remains fixed; a separately locked descriptive condition
only completes the rectangular six-dataset target-model tables.
