\section{Risk alignment through mask geometry}
\label{sec:theory}

\noindent\textbf{Three distinct layers.}
The analysis separates (i) \textbf{cost--geometry compatibility}, which asks
how sensitively \(L(Y,A)\) changes when \(Y\) moves in a chosen mask geometry;
(ii) \textbf{working-posterior adequacy}, which asks whether the declared
\(Q_{p(x)}\) is close to the unknown \(P_x\) in that geometry; and
(iii) \textbf{numerical accuracy}, which asks how well a finite rule integrates
the expectation under \(Q_{p(x)}\). A Lipschitz transport theorem establishes
the first layer, not the second; exact integration removes the third layer,
not posterior discrepancy.

\noindent\textbf{Flagship instance: full surface Hausdorff.}
Let \(d_{H,x}\) be the full Hausdorff distance between inner digital surfaces,
normalized by the image diagonal and extended with zero empty--empty cost and
unit one-sided-empty cost. It is a metric on the resulting surface-set space
and a pseudometric on filled masks. The deployment cost is
\(L_x^{\mathrm{HD}}(A,B)=d_{H,x}(A,B)\in[0,1]\). The reverse triangle
inequality gives
\begin{equation*}
  |L_x^{\mathrm{HD}}(Y,A)-L_x^{\mathrm{HD}}(Y',A)|
  \leq d_{H,x}(Y,Y'),
\end{equation*}
and the general transport lemma in \Cref{app:theory} therefore yields
\begin{equation*}
  \Delta_{L^{\mathrm{HD}},\Yhat}(x)
  \leq W_{1,d_{H,x}}(P_x,Q_{p(x)})
  \leq \TV(P_x,Q_{p(x)}).
\end{equation*}
This is the \textbf{main metric-compatibility guarantee}. Under \(Q_p\), one
scalar level sweeps a coordinated candidate surface; HD records its most
displaced point. The shared-threshold probe and deployment cost therefore use
compatible boundary geometries. The result is conditional: it controls how
posterior discrepancy propagates if \(P_x\) and \(Q_{p(x)}\) are close in
\(d_{H,x}\), but does not establish that closeness. Combining it with
\Cref{sec:framework} controls HD score error and then HD AURC regret.

\noindent\textbf{Robust extension: pooled HD95.}
Replacing the maximum pooled surface distance by its 95th percentile suppresses
a small extreme tail. The resulting normalized penalized HD95 cost remains in
\([0,1]\), so the threshold integral, total-error decomposition, scalar cost-law
transport bound, and AURC theorem all apply. It is not a mask metric, however:
percentile costs can jump when the fraction of distant surface points crosses
five percent and need not satisfy a triangle inequality. Consequently HD95 is
a \textbf{practical robust extension} of the boundary experiment, not a corollary of the
constant-one HD Wasserstein theorem. \Cref{app:rem:hd95-nonmetric} gives an explicit
quantile-jump construction.

\noindent\textbf{Regional contrast: Dice.}
For binary masks, define \(\Dice(A,B)=2|A\cap B|/(|A|+|B|)\),
with \(\Dice(\varnothing,\varnothing)=1\) and Dice zero when exactly one mask
is empty. The deployment cost \(L_{\Dice}=1-\Dice\) lies in \([0,1]\). For Jaccard
distance \(d_J\), \Cref{app:theory} proves
\begin{equation*}
  \Delta_{L_{\Dice},\Yhat}(x)
  \leq\min\{\TV(P_x,Q_{p(x)}),2W_{1,d_J}(P_x,Q_{p(x)})\}.
\end{equation*}
Thus Dice and HD instantiate the \textbf{same Lipschitz--Wasserstein template} under
different, noninterchangeable mask geometries. Dice does not, however, make the
comonotone level-set coupling privileged; coordinated surface sweeps merely
provide a clearer modeling motivation for HD than for Dice.

\Cref{eq:quad} estimates expected Dice under \(Q_p\). Because level
sets change only at probability knots, sorting tied groups gives the exact
integral in \(O(N\log N)\) time. The algorithm in \Cref{app:theory}
removes Dice quadrature error; midpoint integration remains the common estimator
used for controlled cross-risk comparison.

\noindent\textbf{Why SDC and level-set Dice can be close.}
When its denominator is positive, SDC is
\begin{equation}
  \SDC(p,\Yhat)
  =\frac{2\sum_i p_i\hat y_i}{\sum_i p_i+\sum_i\hat y_i},
  \qquad \hat y_i=\mathbf 1\{i\in\Yhat\}.
  \label{eq:sdc}
\end{equation}
For any mask law \(Q\) with marginals \(\Ex_QY_i=p_i\), let
\(N(Y)=2\sum_iY_i\hat y_i\) and
\(D(Y)=\sum_iY_i+\sum_i\hat y_i\). Then
\(\SDC=\Ex_QN/\Ex_QD\), whereas posterior Dice is \(\Ex_Q[N/D]\); they are
not generally equal. For a nonempty action of size \(m\), however,
\begin{equation*}
  \left|\Ex_Q L_{\Dice}(Y,\Yhat)-[1-\SDC(p,\Yhat)]\right|
  \leq \frac{2}{m}\sqrt{\Var_Q |Y\cap\Yhat|}
       +\frac{1}{2m}\sqrt{\Var_Q |Y\setminus\Yhat|}.
\end{equation*}
SDC is therefore the \textbf{count-mean plug-in approximation}: for a fixed
action, spatial coupling affects posterior Dice only through the
\emph{joint} overlap/outside-count law, and its deviation from SDC is
controlled by count dispersion. The \(1/m\) factors can make this explanatory
bound loose for small predicted objects or diffuse outside probability. The
empty action is treated separately because Dice then reduces to the event
\(\{Y=\varnothing\}\) and the displayed bound is undefined. Published SDC
theory derives near-optimality for the ideal Dice confidence under
conditionally independent Bernoulli labels with exact marginals, and then
studies plug-in marginal estimates \citep{sdc}; our level-set score is exact
under the different, comonotone \(Q_p\). Neither assumption dominates the
other.

\noindent\textbf{A marginal-preserving coupling extension.}
To test whether Dice needs non-nested regional variation, we embed the global
level-set law in a spatial Gaussian-copula family. Let
\(\alpha,\beta\geq0\) with \(\alpha+\beta\leq1\). Let \(G\), the field \(U\),
and the iid \(\varepsilon_i\)'s be mutually independent, with
\(G,\varepsilon_i,U_i\sim\mathcal N(0,1)\) marginally. We sample
\begin{equation*}
  Z_i=\sqrt{\alpha}G+\sqrt{\beta}U_i+\sqrt{1-\alpha-\beta}\,\varepsilon_i,
  \qquad Y_i=\mathbf 1\{Z_i\leq\Phi^{-1}(p_i)\}.
\end{equation*}
Every \(Z_i\) is standard normal, hence \(\Pr(Y_i=1)=p_i\) exactly. The family
connects the global shared-threshold endpoint \((\alpha,\beta)=(1,0)\), smooth
non-nested masks, and independent Bernoulli sampling at \((0,0)\). We implement
\(U\) by drawing iid Gaussian coefficients on a coarse coordinate grid whose
knot spacing is \(kD_x\), bilinearly interpolating with image corners aligned,
and dividing each pixel by the analytic interpolation standard deviation.
The result is Gaussian with unit marginal variance; its correlation is
generally nonstationary within coordinate cells. This avoids a dense covariance
matrix. It is an \textbf{experimental coupling-sensitivity extension}, not a
recovered true posterior: it changes dependence while preserving every pixel
marginal, introduces spatial-scale choices and Monte Carlo error, and cannot
identify which coupling generated the labels.
