\section{Experimental protocol}
\label{sec:experiments}

\subsection{Native binary tasks and fixed cohorts}

We evaluate native binary segmentation so that each image contributes exactly
one foreground task and one row to every risk--coverage analysis. We do not
construct ground-truth-selected one-vs-rest cohorts. Every held-out image is
retained, including any image with an empty reference foreground.

\paragraph{Oxford-IIIT Pet.}
We use the official trainval/test split \citep{parkhi2012cats}. Trimap values
1 (pet) and 3 (border) are foreground and value 2 is background. This
predeclared border-inclusive mapping yields a total binary label for every
pixel; no annotation-derived ignore region enters either confidence or risk.
The held-out cohort contains 3,669 images per model condition.

\paragraph{Kvasir-SEG.}
The release contains 1,000 paired polyp images and masks but no official
train/test split \citep{jha2020kvasir}. We rank paired file stems by
\((\operatorname{SHA256}(\text{stem}),\text{stem})\), assign the first
\(\lfloor0.8N\rfloor=800\) to training and the remaining 200 to testing, and order
samples within each split by stem. This deterministic rule is fixed before
model fitting. Because the nominally binary masks are JPEG-compressed, we map
a pixel to foreground when its maximum channel value is at least 128; this
fixed threshold separates the modes near 0 and 255 without treating small
compression residuals as foreground.

\paragraph{FIVES.}
We use the official 600-image training and 200-image test partitions for
retinal-vessel segmentation \citep{jin2022fives}. Any nonzero mask value is
foreground. All dataset pairings and split memberships are validated by stable
sample identifiers before analysis; each model condition therefore contributes
200 held-out rows.

For every model condition, fitting and any model selection use only the
declared training partition. The held-out manifest records the model condition,
checkpoint identity and hash when applicable, split, image identifiers, and
complete confidence/risk schema.
The main analysis requires all ten predeclared dataset--condition pairs defined
below; a missing run makes the analysis incomplete rather than silently changing
the cohort. Model conditions are never selected using test AURC.

\subsection{Prediction models and training}

The primary matrix contains three fixed prediction conditions on each dataset:
\textsc{CLIPSeg-General}, \textsc{CLIPSeg-Target}, and
\textsc{DeepLabV3-Target}.  Within a condition, every confidence method receives
the identical foreground probability map, so confidence comparisons do not
conflate ranking with segmentation-model quality.  A tenth,
\textsc{DeepLabV3-External} condition is declared for Pet only, where the
external vocabulary has an explicit cat/dog mapping.
The four Pet and three conditions on each remaining dataset produce 15,876
condition-specific rows in total.

\paragraph{CLIPSeg.}
We use the public \texttt{CIDAS/clipseg-rd64-refined} checkpoint
\citep{clipseg} at its \(352\times352\) input resolution.  The fixed text prompts
are \texttt{cat} and \texttt{dog} for Pet (combined by a pixelwise maximum),
\texttt{polyp} for Kvasir-SEG, and \texttt{retinal blood vessels} for FIVES.
\textsc{CLIPSeg-General} uses the released weights without access to a target
training image.  For \textsc{CLIPSeg-Target}, we freeze both CLIP encoders and
optimize the segmentation decoder for 40 epochs using binary cross-entropy,
AdamW, learning rate \(10^{-4}\), weight decay \(10^{-4}\), and batch size 32.

\paragraph{DeepLabV3.}
\textsc{DeepLabV3-Target} is DeepLabV3 with a ResNet-50 backbone, initialized
from the torchvision COCO-with-VOC-labels checkpoint \citep{deeplabv3}.  We
replace its main and auxiliary classifiers by two-class heads and fine-tune the
entire network at \(512\times512\) for 40 epochs.  The objective is main
cross-entropy plus \(0.4\) times auxiliary cross-entropy; optimization uses SGD
with learning rate \(5\times10^{-3}\), momentum \(0.9\), weight decay
\(10^{-4}\), and batch size 16.  \textsc{DeepLabV3-External} maps the released
checkpoint's cat and dog probabilities to Pet foreground without target
adaptation.  We do not construct corresponding external conditions for
Kvasir-SEG or FIVES: polyps and retinal vessels are absent from that vocabulary,
so those conditions would not be comparable zero-shot predictors.

Both target-adapted models use a polynomial per-step learning-rate multiplier
\((1-s/S)^{0.9}\), gradient-norm clipping at 1, and seed 0.  Training applies a
shared random scale in \([0.5,2]\), a model-resolution crop (with padded mask
pixels excluded from the loss), and a horizontal flip with probability \(0.5\).
We use the final epoch checkpoint under this fixed schedule; no held-out test
label chooses an epoch, hyperparameter, or condition.  Training uses bfloat16
autocasting when a CUDA device is available.  The run manifest retains target
checkpoint and training-configuration hashes when applicable, evaluator-source
and sample-list hashes, package versions, and device metadata.

\subsection{One action, two losses}

The foreground probability map is resized bilinearly to native image resolution
before thresholding or scoring; masks remain at their native resolution.  Model
inputs are resized to the square resolution stated above with antialiasing.  The
deployed action is fixed throughout:
\begin{equation}
  \Yhat(x)=\{i:p_i(x)\geq\gamma\},
  \qquad \gamma=0.5.
  \label{eq:experimental-action}
\end{equation}
Changing a quadrature threshold never changes this action.

The overlap loss is foreground Dice loss,
\(L_{\Dice}=1-\Dice\), with loss zero for empty--empty and one for a
one-sided empty pair. The boundary loss is normalized penalized HD95,
\(L_{\HDF}\): for two nonempty masks we pool surface-to-surface distances in
both directions, take their 95th percentile, and divide by the native image
diagonal. Empty--empty has loss zero and a one-sided empty pair has loss one.
This places both primary risks in \([0,1]\) and prevents undefined boundary
cases from silently changing the cohort. Raw HD95 in pixels is retained only
as an auxiliary descriptive quantity \citep{karimi2020}.

\subsection{Confidence scores}

For each loss \(L\in\{L_{\Dice},L_{\HDF}\}\), we compute the midpoint score
\begin{equation}
  C_{L,M}(x)
  =-\frac1M\sum_{m=1}^M
  L\!\left(
    \{i:p_i(x)\geq(m-1/2)/M\},
    \Yhat(x)
  \right).
  \label{eq:experimental-score}
\end{equation}
\(M=32\) is the primary estimator. \(M\in\{2,8\}\) forms a predeclared
quadrature-resolution ablation; only node count changes. Dice- and nHD95-indexed
scores use the same action, nodes, weights, spatial resolution, and empty-mask
conventions.

We compare against SDC \citep{sdc}, mean pixelwise maximum probability, and
negative mean binary entropy. For the SDC baseline, the positive-denominator
formula is Equation~\eqref{eq:sdc}; when both the probability mass and hard
mask are empty, we use the published baseline convention of confidence zero.
No confidence score observes the reference mask.

\subsection{Selective evaluation and inference}

For each dataset--condition pair and each risk
\(L\in\{L_{\Dice},L_{\HDF}\}\), we report the exact expected empirical AURC
under uniform random ordering within every exact confidence tie. This makes the
result invariant to input row order. The paper tables report raw AURC and
normalized excess AURC; the machine-readable supplementary analysis additionally
records oracle AURC, random-order AURC, and excess AURC. We define
\begin{equation}
  \operatorname{nAURC}
  =
  \frac{\AURC-\AURC_{\mathrm{oracle}}}
       {\AURC_{\mathrm{random}}-\AURC_{\mathrm{oracle}}},
  \label{eq:normalized-aurc}
\end{equation}
when the denominator is nonzero. Lower is better; zero is oracle ordering and
one is random ordering.

The primary metric-alignment comparison is Dice-\(M32\) versus nHD95-\(M32\)
under both Dice and nHD95 risk. We report the paired AURC difference with a
95\% equal-tail percentile interval from \(B=10{,}000\) paired image-cluster
bootstrap resamples \citep{efron1993}. The same resampled AURC differences
\(\Delta_b^*\) define a two-sided percentile-bootstrap tail probability,
\begin{equation}
  p_{\mathrm{boot}}
  =\min\!\left\{1,\;2\min\!\left(
    \frac{1+\sum_{b=1}^{B}\mathbf{1}\{\Delta_b^*\leq0\}}{B+1},
    \frac{1+\sum_{b=1}^{B}\mathbf{1}\{\Delta_b^*\geq0\}}{B+1}
  \right)\right\}.
  \label{eq:bootstrap-p}
\end{equation}
Thus the effect, interval, and raw \(p\)-value target the same test-image
sampling estimand and are invariant to strictly increasing reparameterizations
of either confidence score. We apply Holm's step-down family-wise-error
adjustment to these raw \(p\)-values across the 20 predeclared comparisons
(ten dataset--condition pairs and two risks) \citep{holm1979}. The reported
percentile intervals are pointwise, not multiplicity-adjusted. Baselines and
\(M\)-ladder comparisons are descriptive and excluded from that family.
All seeds, hashes, row counts, and generated artifacts are written by the
analysis pipeline.
