\section{Results}
\label{sec:results}

Geometry can change the safest ordering; matched scores need not win. We report
AURC-derived scalars \(\times100\) (not nAURC); negative contrasts favor the left.

\IfFileExists{Tables/results_complete.tex}{%
  \IfFileExists{Tables/main_results.tex}{%
    \input{Tables/main_results.tex}
  }{%
    \paragraph{Rendering guard.} The final analysis is incomplete.
  }
}{%
  \paragraph{Rendering guard.} The final analysis is incomplete.
}

\paragraph{Region overlap to worst-case boundary.}
nHD-M32 lowers nHD-risk AURC relative to Dice-M32 in all ten target conditions,
with all ten pointwise intervals excluding zero. Under Dice risk, Dice-M32 is
lower in eight, but only three intervals exclude zero. The strongest separation
is therefore worst-boundary risk, not a symmetric same-name advantage.

\paragraph{Worst-case to robust boundary.}
nHD-M32 is lower than nHD95-M32 in nine target conditions under nHD risk;
nHD95-M32 is lower in eight under nHD95 risk. Only four and two matched-direction
intervals exclude zero. The smaller, condition-dependent gains do not transfer
the nHD Wasserstein corollary to nHD95.

\paragraph{Training-seed sensitivity.}
The prespecified training-sensitivity rule is met only for nHD--nHD95 under
nHD95 risk: five of ten conditions reverse across seeds and seed 0 disagrees
with the majority in three; Appendix~\ref{app:additional-results} reports the
checkpoint-level values. We call this training-sensitive, not seed inference.

\paragraph{Baselines and numerical fidelity.}
A boundary-indexed score has the lowest raw AURC in nine of ten target
conditions under each boundary risk; SDC on FIVES/CLIPSeg-Target is the
exception. SDC and foreground entropy remain competitive for Dice. Dice-M32 and
Dice-Exact have \(\rho\geq.9992\), \(\tau_b\geq.9791\), and at most .363 displayed
Dice-risk AURC difference. For nHD/nHD95, M32--M128 \(\rho\) is
.973--.998/.951--.996 and the largest displayed gaps are .456/.324.

\paragraph{Mechanism clues.}
In nine target conditions other than the failed FIVES/CLIPSeg-Target cell,
Dice-M32 and foreground entropy have score--score Spearman correlation
.916--.960; the exceptional cell is .535. Replacing Dice-M32 by Dice-Exact
changes its association with realized Dice safety by at most .004. By contrast,
nHD-M32 has higher association with realized nHD safety than Dice-M32 in all ten
target conditions, with differences .179--.432. This is consistent with Dice
redundancy and additional spatial-extreme information in nHD indexing, but it
does not establish that real annotation variability is nested or identify the
cause of Dice errors. Appendix~\ref{app:additional-results} reports the complete
diagnostic table; a fixed-action ensemble-posterior extension directly tests the
support explanation at a higher inference budget.

An action-two-block Dice posterior passes its frozen 8/10 directional gate, but
mean/median AURC\(\times100\) gains are only .0195/.0020 and its rank correlation
with Dice-M32 is at least .99936; count decoupling does not remove the redundancy.
A separately predeclared component--grid ladder reaches only 6/10 wins over
two-block for every component variant, so none passes its gate. Action-aligned
components beat their equal-cardinality grids in 9/10 conditions, but by only
.0635/.0446 mean/median AURC\(\times100\); proposal variants are less consistent,
and no component construction beats SDC or foreground entropy in more than
6/10 conditions. Spatial grouping can change the count law, but does not yield
a robust Dice-ranking advantage.
The higher-budget three-checkpoint Ensemble-Q fails its Dice gate: LevelSet-Q is
lower in 9/10 conditions and its mean advantage is 2.763 AURC\(\times100\).
LevelSet-Q nHD is lower in all ten, by 3.820 on average; raw checkpoint
disagreement is therefore not an annotation-posterior substitute.

\paragraph{Stress tests.}
Known-posterior tests show TV near one while Dice score error remains
.0135--.0393; cardinality checks find zero support for up to 11.5\% of FIVES and
15.4\% of ISIC references. Both boundary macro contrasts reverse at
\(\gamma=.3\), whereas the Dice--nHD directions persist.

Appendix~\ref{app:additional-results} reports all AURCs, 64 pointwise intervals,
cross-loss panels, agreement, curves, and stress tests.
