\section{Limitations and Conclusion}
\label{sec:limitations}
\label{sec:limitations-conclusion}

The shared-threshold law is \textbf{a boundary-motion model, not an identified posterior}.
More quadrature nodes reduce numerical error, not
posterior discrepancy. The nHD theorem states a sufficient surface-geometric
condition; one label per image cannot verify it, and observed support failures
show where it can break.

Full nHD is sensitive to one extreme surface point. Pooled nHD95 is more robust
but nonmetric and empirically sensitive to checkpoint and action threshold.
Surface extraction, spacing, normalization, pooling, and empty-mask penalties
must therefore be fixed as deployment choices. Our five-dataset primary study
and separate SegFormer/DUTS extension broaden architecture and domain coverage,
but one primary seed and an image bootstrap do not cover site,
multi-rater, patient-cluster, or training uncertainty. Learned detectors,
calibrated annotation posteriors, multiclass actions, and absolute risk
calibration remain future work; \Cref{app:limitations} gives details.

\noindent\textbf{Conclusion.}
\label{sec:conclusion}
Selective segmentation begins by \textbf{naming the failure that should trigger
abstention}. Loss indexing supplies the target, but the working posterior must
also encode how that failure occurs. Shared thresholds model coherent boundary
motion, and normalized full Hausdorff loss yields the clearest theory and
empirical ranking across both the primary and architecture--domain studies.
nHD95 exposes the value and sensitivity of a robust endpoint;
Dice shows parity between level-set and count-oriented SDC across the target
conditions, consistent with the count-dispersion analysis.
\textbf{Confidence is not universal or metric-free}: it is a declared model of how a
fixed deployed mask may be wrong.
