\section{Expanded Benchmark Protocol}
\label{app:benchmark-protocol}

This appendix freezes the expanded five-dataset protocol and comparator
definitions independently of their observed AURC. Its purpose is to make every
row reconstructible, including cases in which a predicted foreground is empty.

\subsection{Native-binary cohorts}

\paragraph{Oxford-IIIT Pet.}
We use the official trainval/test split: 3,680 released training examples and
3,669 test examples \citep{parkhi2012cats}. Trimap values 1 (pet) and 3 (border)
are foreground and value 2 is background. This border-inclusive mapping is
declared before evaluation and leaves no annotation-derived ignore pixels.

\paragraph{Kvasir-SEG.}
The release contains 1,000 paired polyp images and masks but no official split
\citep{jha2020kvasir}. We rank stems by
\((\operatorname{SHA256}(\text{stem}),\text{stem})\), assign the first 800 to
training and the remaining 200 to testing, and order each split by stem. For its
JPEG-compressed masks, a pixel is foreground iff the maximum decoded channel is
at least 128; this rejects small compression residuals near zero.

\paragraph{FIVES.}
We use the official 600-image training and 200-image test partitions
\citep{jin2022fives}. A pixel is foreground iff its mask value is nonzero.

\paragraph{ISIC 2018 Task 1.}
We use the official 2,594-image training and 1,000-image test releases for lesion
boundary segmentation \citep{codella2019isic,tschandl2018ham10000,
isic2018task1}. Each losslessly encoded mask denotes the primary skin lesion
with values 0 and 255; foreground is therefore the nonzero set. Image and mask
identifiers are joined by the official \texttt{ISIC\_id} stem rather than by
directory order.

\paragraph{TN3K.}
We use the released split of 2,879 training and 614 test ultrasound images
\citep{gong2021multitask,gong2023thyroid}. The source contains 3,493 images from
2,421 patients and permits multiple disconnected nodules in one binary mask. As
the released masks are JPEG-compressed, foreground again means maximum decoded
channel value at least 128.

For all five datasets, image--mask dimensions, one-to-one stem pairing, split
membership, and stable sample identifiers are checked before inference. The
fixed CLIPSeg prompts are \texttt{cat}/\texttt{dog}, \texttt{polyp},
\texttt{retinal blood vessels}, \texttt{skin lesion}, and
\texttt{thyroid nodule}, respectively. Scores and risks are computed after
bilinear probability-map resizing to the native image grid; masks are never
interpolated for evaluation.

\subsection{Prediction models and frozen training schedules}

We use \texttt{CIDAS/clipseg-rd64-refined} \citep{clipseg} at
\(352\times352\). \textsc{CLIPSeg-General} uses its released weights;
\textsc{CLIPSeg-Target} freezes both CLIP encoders and optimizes the decoder
for 40 epochs with binary cross-entropy, AdamW, learning rate \(10^{-4}\),
weight decay \(10^{-4}\), and batch size 32.

\textsc{DeepLabV3-Target} is a ResNet-50 DeepLabV3 initialized from the
torchvision COCO-with-VOC-labels checkpoint \citep{deeplabv3}. Both classifiers
are replaced by two-class heads and fine-tuned at \(512\times512\) for 40
epochs. The objective is main cross-entropy plus \(0.4\) times auxiliary
cross-entropy; SGD uses learning rate \(5\times10^{-3}\), momentum \(0.9\),
weight decay \(10^{-4}\), and batch size 16. The Pet-only
\textsc{DeepLabV3-External} condition maps the released cat and dog
probabilities to foreground.

Both target-adapted schedules use the per-step multiplier
\((1-s/S)^{0.9}\), gradient clipping at 1, and seed 0. Augmentation applies a
shared random scale in \([0.5,2]\), a model-resolution crop with padded pixels
excluded from loss, and a horizontal flip with probability \(0.5\). We retain
the final epoch under this predeclared schedule. Frozen manifests record the
checkpoint, source, cohort, configuration, package, and device hashes or
identifiers.

\subsection{Single-map entropy comparators}

All entropy comparators begin with the same foreground probabilities
\(p=(p_i)_{i\in\Omega}\), deployed threshold \(\gamma\), and hard foreground
\(F_\gamma=\{i:p_i\geq\gamma\}\). Let \(n=|\Omega|\), \(k=|F_\gamma|\), and
\[
  u_i=-p_i\log_2p_i-(1-p_i)\log_2(1-p_i),
  \qquad 0\log_2 0:=0.
\]
Thus \(u_i\in[0,1]\). Every reported scalar below is oriented as confidence:
higher is better.

\paragraph{QFR-Entropy.}
The foreground-ratio aggregator sets \(q_{\mathrm{FG}}=k/n\), computes
\(\tau=Q_{1-q_{\mathrm{FG}}}(u)\) with NumPy's default linear quantile, and
uses the inclusive cutoff from the reference implementation
\citep{guarino2026aggregation}:
\[
 C_{\mathrm{QFR}}
 =-\operatorname{mean}\{u_i:u_i\geq\tau\}.
\]
All cutoff ties are therefore retained, so their realized fraction can exceed
\(q_{\mathrm{FG}}\). If \(k=0\), then \(\tau=\max_i u_i\) and the score is
\(-\max_i u_i\); the selected set is never empty. This convention reproduces
the public implementation exactly rather than silently replacing it by a
top-\(k\) rule.

\paragraph{PLM-10/PLA-10-Entropy.}
For the set \(\mathcal W_{10}\) of all valid, unpadded \(10\times10\) windows on
the native grid, we use
\[
 C_{\mathrm{PLM10}}
 =-\max_{w\in\mathcal W_{10}}\frac{1}{|w|}\sum_{i\in w}u_i.
\]
PLM is commonly written as a maximum patch average
\citep{guarino2026aggregation}; PLA-10 is written as the negative maximum patch
sum \citep{sdc}. Since every valid window here has 100 pixels, the two scores
differ only by the positive constant 100 and therefore induce exactly the same
ranking and AURC. We report one row rather than counting rank-equivalent methods
twice.

\paragraph{MMMC-Entropy.}
Following the entropy form compared by \citet{sdc} and originating with
\citet{holder2021uncertainty},
\[
 C_{\mathrm{MMMC}}=
 \begin{cases}
 -\bigl(\operatorname{median}_i u_i+\min_i u_i\bigr)/\max_i u_i,
   & \max_i u_i>0,\\
 0, & \max_i u_i=0.
 \end{cases}
\]
The second branch fixes the otherwise undefined all-zero-entropy case.

\paragraph{Foreground entropy.}
The action-restricted score is
\[
 C_{\mathrm{FGEnt}}=
 \begin{cases}
  -k^{-1}\sum_{i\in F_\gamma}u_i, & k>0,\\
  -1, & k=0.
 \end{cases}
\]
The empty-action value is the bottom of the nonempty score range, rather than a
global-entropy fallback. Mean negative entropy
\(-n^{-1}\sum_i u_i\), mean maximum probability
\(n^{-1}\sum_i\max(p_i,1-p_i)\), and SDC from
Equation~\eqref{eq:sdc} complete the reference panel. SDC retains its published
zero confidence when both its probability mass and hard mask have zero mass.

\subsection{Matched information budget}

Every primary confidence method is a deterministic function of the same single
foreground probability map and fixed deployed action. It receives no reference
mask, internal feature, validation-label fit, test-set rank transform, extra
forward pass, or method-specific calibration. QFR uses the current image's
predicted foreground ratio, PLM/PLA fixes its patch width at 10, and no comparator
is selected by test AURC.

Monte Carlo dropout, test-time augmentation, and deep ensembles produce multiple
maps and therefore change both information and inference cost. Learned failure
predictors additionally use internal features and a labeled fitting set. These
are valuable but answer a different-budget question; we exclude them from the
matched primary table and treat them only as extensions discussed in the
limitations. This isolates the paper's question: given one fixed probability
map, does changing only the image-level confidence functional improve selective
ranking?

\subsection{Loss, numerical, and inference conventions}

For a mask \(A\), the inner digital surface is
\(S(A)=A\setminus\operatorname{erosion}(A)\), using SciPy
\texttt{binary\_erosion}'s default structuring element and border handling;
nearest-surface distances use native-grid Euclidean distance transforms.
For nonempty masks, the native-pixel boundary auxiliaries are
\(\max D(A,B)\) and \(Q_{0.95}(D(A,B))\). A one-sided empty pair receives the
image diagonal. Their normalized primary counterparts divide by that diagonal
and use loss one for a one-sided empty pair; empty--empty loss is zero. The
percentile is NumPy's default linear quantile after pooling both directed
surface-distance arrays. This fixes the otherwise consequential pooling,
interpolation, and empty-mask choices.

For Dice-Exact versus each \(M\in\{2,8,32\}\) midpoint score, we retain the
mean, median, 95th percentile, and maximum absolute per-image score error,
Spearman correlation, Kendall's \(\tau_b\), and exact-match fraction.
Correlations use average ranks and are undefined when either score is constant.
Separately, each locked probability map yields pixel-weighted Brier score,
15-bin ECE, truth- and prediction-empty rates, mean distinct-mask count along
the \(M=32\) ladder, the fraction of unchanged adjacent threshold pairs, and
the mean adjacent changed-pixel fraction. These held-out-label diagnostics are
descriptive only: they do not tune a score and cannot identify the joint mask
posterior.

Empirical AURC is the exact expectation under uniform random ordering within
every exact confidence tie. We also retain oracle AURC, random-order AURC, and
normalized excess AURC whenever its denominator is nonzero. Each predeclared
adjacent-geometry contrast uses the same 10,000 paired image-cluster bootstrap
resamples for both scores and the observed risk \citep{efron1993}. Equal-tail
intervals are pointwise. Holm's step-down transform \citep{holm1979} is applied
separately to the 40-comparison Pet/Kvasir/FIVES core family and the
24-comparison ISIC/TN3K extension family. Because the inputs are approximate
bootstrap confidence-curve tail areas rather than exact null-calibrated
\(p\)-values, we report them as multiplicity-aware evidence summaries and do
not claim exact finite-sample family-wise error control.
